---
title: "Prelim Analysis & Visualisations"
author: "Irene & Beatrice"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
#timeseries for TRAIN
library(ggplot2)
library(dplyr)
library(tidyr)

# Read the dataset
pv_train_od_202312 <- read_csv("../data/origin_destination_train_202312.csv")

pv_train_od_202312 <- pv_train_od_202312 %>%
  mutate(Location = ifelse(ORIGIN_PT_CODE < DESTINATION_PT_CODE, ORIGIN_PT_CODE, DESTINATION_PT_CODE))

# Grouping data by TIME_PER_HOUR and location, and summing up the total trips
grouped_data <- pv_train_od_202312 %>%
  group_by(TIME_PER_HOUR, Location) %>%
  summarise(Total_Trips = sum(TOTAL_TRIPS))

# Sorting the total trips per location in descending order and selecting top 20 locations
top_20_locations <- grouped_data %>%
  group_by(Location) %>%
  summarise(Total_Trips = sum(Total_Trips)) %>%
  top_n(20, Total_Trips)

# Filter the top 20 locations
grouped_data <- grouped_data %>%
  filter(Location %in% top_20_locations$Location)

# Plotting the time series graph
ggplot(grouped_data, aes(x = TIME_PER_HOUR, y = Total_Trips, color = as.factor(Location))) +
  geom_line() +
  labs(title = "Total Trips vs. Time for Top 20 Locations",
       x = "Time (Hour)",
       y = "Total Trips",
       color = "Location") +
  theme_minimal() +
  theme(legend.position = "top")
```

```{r}
#time series, grp by weekends, weekdays (TRAIN)

pv_train_od_202312 <- read_csv("../data/origin_destination_train_202312.csv")
library(ggplot2)
library(dplyr)
library(tidyr)

pv_train_od_202312 <- 
pv_train_od_202312 %>%
  mutate(Location = ifelse(ORIGIN_PT_CODE < DESTINATION_PT_CODE, ORIGIN_PT_CODE, DESTINATION_PT_CODE))

# Grouping data by TIME_PER_HOUR, day type, and location, and summing up the total trips
grouped_data <- 
pv_train_od_202312 %>%
  group_by(TIME_PER_HOUR, DAY_TYPE, Location) %>%
  summarise(Total_Trips = sum(TOTAL_TRIPS))

# Sorting the total trips per location in descending order and selecting top 20 locations
top_20_locations <- grouped_data %>%
  group_by(Location) %>%
  summarise(Total_Trips = sum(Total_Trips)) %>%
  top_n(20, Total_Trips)

# Filter the top 20 locations
grouped_data <- grouped_data %>%
  filter(Location %in% top_20_locations$Location)

# Plotting the time series graph
ggplot(grouped_data, aes(x = TIME_PER_HOUR, y = Total_Trips, color = as.factor(Location))) +
  geom_line() +
  labs(title = "Total Trips vs. Time for Top 20 Locations",
       x = "Time (Hour)",
       y = "Total Trips",
       color = "Location") +
  facet_wrap(~DAY_TYPE) +  # Adding facet_wrap for weekdays and weekends
  theme_minimal() +
  theme(legend.position = "top")
```

```{r}

pv_bus_od_202402 <- read.csv("../data/origin_destination_bus_202402.csv")

pv_bus_od_202402 <- pv_bus_od_202402 %>%
  mutate(Location = ifelse(ORIGIN_PT_CODE < DESTINATION_PT_CODE, ORIGIN_PT_CODE, DESTINATION_PT_CODE))

# Grouping data by TIME_PER_HOUR, day type, and location, and summing up the total trips
grouped_databus <- pv_bus_od_202402 %>%
  group_by(TIME_PER_HOUR, DAY_TYPE, Location) %>%
  summarise(Total_Trips = sum(TOTAL_TRIPS))

# Sorting the total trips per location in descending order and selecting top 10 locations
top_30_locations_bus <- grouped_databus %>%
  group_by(Location) %>%
  summarise(Total_Trips = sum(Total_Trips)) %>%
  top_n(30, Total_Trips)

# Filter the top 30 locations
grouped_databus1 <- grouped_databus %>%
  filter(Location %in% top_30_locations_bus$Location)

# Plotting the time series graph
ggplot(grouped_databus1, aes(x = TIME_PER_HOUR, y = Total_Trips, color = as.factor(Location))) +
  geom_line() +
  labs(title = "Total Trips vs. Time for Top 10 Locations",
       x = "Time (Hour)",
       y = "Total Trips",
       color = "Location") +
  facet_wrap(~DAY_TYPE) +  # Adding facet_wrap for weekdays and weekends
  theme_minimal() +
  theme(legend.position = "top")
```



## Loading in data
```{r}
# Passenger Volume for Buses - By Bus Stops (2023 Dec)
pv_bus_202312 <- read.csv("../data/transport_node_bus_202312.csv")

# Passenger Volume for Buses - By Bus Stops (2024 Feb)
pv_bus_202402 <- read.csv("../data/transport_node_bus_202402.csv")

# Passenger Volume for Buses - By Origin Destination Bus Stops (2023 Dec)
pv_bus_od_202312 <- read.csv("../data/origin_destination_bus_202312.csv")

# Passenger Volume for Buses - By Origin Destination Bus Stops (2024 Feb)
pv_bus_od_202402 <- read.csv("../data/origin_destination_bus_202402.csv")



# Passenger Volume for Trains - By Train Stops (2023 Dec)
pv_train_202312 <- read.csv("../data/transport_node_train_202312.csv")

# Passenger Volume for Trains - By Train Stops (2024 Feb)
pv_train_202402 <- read.csv("../data/transport_node_train_202402.csv")

# Passenger Volume for Trains - By Origin Destination (2023 Dec)
#pv_train_od_202312 <- read.csv("../PV Data [Buses and Trains]/Trains/By OD/origin_destination_train_202312.csv")
pv_train_od_202312 <- read.csv("../data/origin_destination_train_202312.csv")

# Passenger Volume for Trains - By Origin Destination (2024 Feb)
pv_train_od_202402 <- read.csv("../data/origin_destination_train_202402.csv")

# Geography of Island
library(sf)
shape = st_make_valid(st_read("../data/MasterPlan2019SubzoneBoundaryNoSeaGEOJSON.geojson"))
ggplot(data = shape) +
  geom_sf(aes(geometry = geometry), fill = "lightgray", color = "white") +
  labs(title = "Subzone bounaries in Master Plan 2019")


# Train Stations in Singapore (using data from Kaggle)
shape_train = read_csv("../data/TrainStations.csv") %>%
  select(-geometry) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = st_crs(shape))

ggplot() +
  geom_sf(data = shape, aes(geometry = geometry), fill = "steelblue", alpha = 0.3,
          color = "white") + labs(title = "Subzone bounaries in Master
                                  Plan 2019") +
  geom_sf(data = shape_train, aes(geometry = geometry), color = "darkblue") +
  theme_minimal()


# Bus Stops in Singapore
shape_bus <- st_read("../data/BusStopLocation_Jul2023/BusStop.shp")
ggplot() +
  geom_sf(data = shape, aes(geometry = geometry), fill = "steelblue", alpha = 0.3,
          color = "white") + labs(title = "Subzone bounaries in Master
                                  Plan 2019") +
  geom_sf(data = shape_bus, aes(geometry = geometry), color = "darkblue", size=0.8)
```


```{r}
library(plotly)
library(patchwork)

#busstop_in_region: merging shape_bus and shape, to see which region each bus stop belongs to
shape_bus = st_transform(shape_bus, st_crs(shape))
busstop_in_region = st_join(shape_bus, shape, left = F)

#trainstn_in_region: merging shape_train and shape, to see which region each train stn belongs to
shape_train = st_transform(shape_train, st_crs(shape))
trainstn_in_region = st_join(shape_train, shape, left = F)

###pv_bus, 202402
pv_bus_feb = pv_bus_202402 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  group_by(PT_CODE) %>%
  summarise(TOTAL_VOLUME = sum(VOLUME)) %>%
  arrange(desc(TOTAL_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

pv_bus_feb_weekday = pv_bus_202402 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/20) %>% #Feb have 20 weekdays
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

pv_bus_feb_weekend = pv_bus_202402 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKENDS/HOLIDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/9) %>% #Feb have 9 weekends/hol
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

###pv_bus, 202312
pv_bus_dec = pv_bus_202312 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  group_by(PT_CODE) %>%
  summarise(TOTAL_VOLUME = sum(VOLUME)) %>%
  arrange(desc(TOTAL_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

pv_bus_dec_weekday = pv_bus_202312 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/20) %>% #Dec have 20 weekdays
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

pv_bus_dec_weekend = pv_bus_202312 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKENDS/HOLIDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/11) %>% #Dec have 11 weekends/hol
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))
```



## Heatmap of number of trips to regions in Dec23 & Feb24, all day types
```{r}
busstop_in_region_dec = busstop_in_region %>%
  left_join(pv_bus_dec, by = c("BUS_STOP_N" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(TOTAL_VOLUME = sum(TOTAL_VOLUME)) %>%
  select(Name, TOTAL_VOLUME)

busstop_in_region_dec %>%   #top 10 regions that have the highest traffic in the area
  arrange(desc(TOTAL_VOLUME)) %>%
  head(10)

p1 = st_join(shape, busstop_in_region_dec) %>%
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = TOTAL_VOLUME)) +
  scale_fill_continuous(name = "No. of people travelling to region via bus", low = 'lightblue', high = 'darkblue') +
  labs(title = "Heatmap of Passenger Volume in the different regions by bus in Dec") +
  theme_minimal()
p1 = ggplotly(p1)
p1 #p1: heatmap of total volume of passengers taking bus in diff regions, dec, all day types



busstop_in_region_feb = busstop_in_region %>%
  left_join(pv_bus_feb, by = c("BUS_STOP_N" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(TOTAL_VOLUME = sum(TOTAL_VOLUME)) %>%
  select(Name, TOTAL_VOLUME)

busstop_in_region_feb %>%   #top 10 regions that have the highest traffic in the area
  arrange(desc(TOTAL_VOLUME)) %>%
  head(10)

p3 = st_join(shape, busstop_in_region_feb) %>%
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = TOTAL_VOLUME)) +
  scale_fill_continuous(name = "No. of people travelling to region via bus", low = 'lightblue', high = 'darkblue') +
  labs(title = "Heatmap of Passenger Volume in the different regions by bus in Feb") +
  theme_minimal()
p3 = ggplotly(p3)
p3 #p3: heatmap of total volume of passengers taking bus in diff regions, feb, all day types

subplot(p1, p3, shareY=TRUE, shareX = TRUE) %>% #side-by-side plots of heatmaps of PV in Dec and Feb
  layout(xaxis = list(title = "Dec"), xaxis2 = list(title = "Feb"), title = "Heatmap of Passenger Volume in different regions by bus, Dec vs Feb", annotations = list(list(x = 0.2,  y = 1.0,  text = "Dec"),  list(x = 0.8,  y = 1,  text = "Plot 2")))


```


## Lollipop graph comparing weekday and weekend/holiday passenger volume in top 10 regions, in Dec
```{r}
busstop_in_region_wd_dec = busstop_in_region %>%
  left_join(pv_bus_dec_weekday, by = c("BUS_STOP_N" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekday")

busstop_in_region_we_dec = busstop_in_region %>%
  left_join(pv_bus_dec_weekend, by = c("BUS_STOP_N" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekend/holiday")

busstop_in_region_wdVSwe_dec = rbind(busstop_in_region_wd_dec, busstop_in_region_we_dec)
p2 = busstop_in_region_wdVSwe_dec %>%
  ggplot(aes(x=Name ,y=AVG_DAILY_VOLUME)) +
  geom_line(aes(group=Name), color = 'grey') +
  geom_point(aes(color = day_type), alpha = 0.7) +
  labs(title = "Total trips to each region, weekday vs weekend, in Dec 2023", y="No. of trips to", x = "Region code")
p2 = ggplotly(p2)
p2 #p2: comparing passenger volume to diff regions between weekday and weekend/holiday, in the top 10 most crowded regions



```



## Lollipop graph comparing weekday and weekend/holiday passenger volume in top 10 regions, in Feb
```{r}
busstop_in_region_wd_feb = busstop_in_region %>%
  left_join(pv_bus_feb_weekday, by = c("BUS_STOP_N" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekday")

busstop_in_region_we_feb = busstop_in_region %>%
  left_join(pv_bus_feb_weekend, by = c("BUS_STOP_N" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekend/holiday")

busstop_in_region_wdVSwe_feb = rbind(busstop_in_region_wd_feb, busstop_in_region_we_feb)
p4 = busstop_in_region_wdVSwe_feb %>%
  ggplot(aes(x=Name ,y=AVG_DAILY_VOLUME)) +
  geom_line(aes(group=Name), color = 'grey') +
  geom_point(aes(color = day_type), alpha = 0.7) +
  labs(title = "Total trips to each region, weekday vs weekend, in Feb 2023", y="No. of trips to", x = "Region code")
p4 = ggplotly(p4)
p4 #p4: comparing passenger volume to diff regions between weekday and weekend/holiday, in the top 10 most crowded regions
```


```{r}
###pv_train, 202402
pv_train_feb = pv_train_202402 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  group_by(PT_CODE) %>%
  summarise(TOTAL_VOLUME = sum(VOLUME)) %>%
  arrange(desc(TOTAL_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))


pv_train_feb_weekday = pv_train_202402 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/20) %>% #Feb have 20 weekdays
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

pv_train_feb_weekend = pv_train_202402 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKENDS/HOLIDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/9) %>% #Feb have 9 weekends/hol
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

###pv_bus, 202312
pv_train_dec = pv_train_202312 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  group_by(PT_CODE) %>%
  summarise(TOTAL_VOLUME = sum(VOLUME)) %>%
  arrange(desc(TOTAL_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

pv_train_dec_weekday = pv_train_202312 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/20) %>% #Dec have 20 weekdays
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))

pv_train_dec_weekend = pv_train_202312 %>%
  select(-YEAR_MONTH, -TIME_PER_HOUR) %>%
  mutate(DAY_TYPE = as.factor(DAY_TYPE), VOLUME = TOTAL_TAP_IN_VOLUME + TOTAL_TAP_OUT_VOLUME) %>%
  filter(DAY_TYPE == 'WEEKENDS/HOLIDAY') %>%
  group_by(PT_CODE) %>%
  summarise(AVG_DAILY_VOLUME = sum(VOLUME)/11) %>% #Dec have 11 weekends/hol
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  mutate(PT_CODE = as.character(PT_CODE))
```


##Heatmap of number of trips to regions in Dec23 & Feb24, all day types
```{r}
trainstn_in_region_dec = trainstn_in_region %>%
  left_join(pv_train_dec, by = c("STN_NO" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(TOTAL_VOLUME = sum(TOTAL_VOLUME)) %>%
  select(Name, TOTAL_VOLUME)

trainstn_in_region_dec %>%   #top 10 regions that have the highest traffic in the area
  arrange(desc(TOTAL_VOLUME)) %>%
  head(10)

p5 = st_join(shape, trainstn_in_region_dec) %>%
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = TOTAL_VOLUME)) +
  scale_fill_continuous(name = "No. of people travelling to region via train", low = 'lightblue', high = 'darkblue') +
  labs(title = "Heatmap of Passenger Volume in the different regions by train in Dec") +
  theme_minimal() +
  geom_sf(data = shape_train, aes(geometry = geometry), color = "red", size = 0.1)
p5 = ggplotly(p5)
p5 #p5: heatmap of total volume of passengers taking bus in diff regions, dec, all day types


##Heatmap of number of trips to regions in Feb, all day types
trainstn_in_region_feb = trainstn_in_region %>%
  left_join(pv_train_feb, by = c("STN_NO" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(TOTAL_VOLUME = sum(TOTAL_VOLUME)) %>%
  select(Name, TOTAL_VOLUME)

trainstn_in_region_feb %>%   #top 10 regions that have the highest traffic in the area
  arrange(desc(TOTAL_VOLUME)) %>%
  head(10)

p7 = st_join(shape, trainstn_in_region_feb) %>%
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = TOTAL_VOLUME)) +
  scale_fill_continuous(name = "No. of people travelling to region via train", low = 'lightblue', high = 'darkblue') +
  labs(title = "Heatmap of Passenger Volume in the different regions by train in Feb") +
  theme_minimal() +
  geom_sf(data = shape_train, aes(geometry = geometry), color = "red", size = 0.1)
p7 = ggplotly(p7)
p7 #p7: heatmap of total volume of passengers taking bus in diff regions, dec, all day types

subplot(p5, p7, shareY=TRUE, shareX = TRUE) %>% #side-by-side plots of heatmaps of PV in Dec and Feb
  layout(xaxis = list(title = "Dec"), xaxis2 = list(title = "Feb"), title = "Heatmap of Passenger Volume in different regions by train, Dec vs Feb", annotations = list(list(x = 0.2,  y = 1.0,  text = "Dec"),  list(x = 0.8,  y = 1,  text = "Plot 2")))

```

## Lollipop graph comparing weekday and weekend/holiday passenger volume in top 10 regions, in Dec
```{r}
trainstn_in_region_wd_dec = trainstn_in_region %>%
  left_join(pv_train_dec_weekday, by = c("STN_NO" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekday")

trainstn_in_region_we_dec = trainstn_in_region %>%
  left_join(pv_train_dec_weekend, by = c("STN_NO" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekend/holiday")

trainstn_in_region_wdVSwe_dec = rbind(trainstn_in_region_wd_dec, trainstn_in_region_we_dec)
p6 = trainstn_in_region_wdVSwe_dec %>%
  ggplot(aes(x=Name ,y=AVG_DAILY_VOLUME)) +
  geom_line(aes(group=Name), color = 'grey') +
  geom_point(aes(color = day_type), alpha = 0.7) +
  labs(title = "Total trips to each region, weekday vs weekend, in Dec 2023", y="No. of trips to", x = "Region code")
p6 = ggplotly(p6)
p6 #p6: comparing passenger volume to diff regions between weekday and weekend/holiday, in the top 10 most crowded regions



```



## Lollipop graph comparing weekday and weekend/holiday passenger volume in top 10 regions, in Feb
```{r}
trainstn_in_region_wd_feb = trainstn_in_region %>%
  left_join(pv_train_feb_weekday, by = c("STN_NO" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekday")

trainstn_in_region_we_feb = trainstn_in_region %>%
  left_join(pv_train_feb_weekend, by = c("STN_NO" = "PT_CODE")) %>%
  group_by(Name) %>%
  summarise(AVG_DAILY_VOLUME = sum(AVG_DAILY_VOLUME)) %>%
  select(Name, AVG_DAILY_VOLUME) %>%
  arrange(desc(AVG_DAILY_VOLUME)) %>%
  head(10) %>%
  mutate(day_type = "weekend/holiday")

trainstn_in_region_wdVSwe_feb = rbind(trainstn_in_region_wd_feb, trainstn_in_region_we_feb)
p8 = trainstn_in_region_wdVSwe_feb %>%
  ggplot(aes(x=Name ,y=AVG_DAILY_VOLUME)) +
  geom_line(aes(group=Name), color = 'grey') +
  geom_point(aes(color = day_type), alpha = 0.7) +
  labs(title = "Total trips to each region, weekday vs weekend, in Dec 2023", y="No. of trips to", x = "Region code")
p8 = ggplotly(p8)
p8 #p8: comparing passenger volume to diff regions between weekday and weekend/holiday, in the top 10 most crowded regions
```



### Popular Train Routes map
```{r}
popular_routes_train_feb = pv_train_od_202402 %>%
  group_by(ORIGIN_PT_CODE, DESTINATION_PT_CODE) %>%
  summarise(trips = sum(TOTAL_TRIPS)) %>%
  arrange(desc(trips)) %>%
  head(20) %>%
  separate(ORIGIN_PT_CODE, into = c("ORIGIN_PT", "ORIGIN_PT2"), sep = '/') %>%
  separate(DESTINATION_PT_CODE, into = c("DEST_PT", "DEST_PT2"), sep = '/')

  
shape_train_sep = shape_train %>%
  separate(STN_NO, into = c("STN_NO", "STN_NO2"), sep = '/')
pop_train_routes = popular_routes_train_feb %>%
  left_join(shape_train_sep, by = c("ORIGIN_PT" = "STN_NO")) %>%
  select(ORIGIN_PT, ORIGIN_PT2, DEST_PT, DEST_PT2, trips, STN_NAME, geometry) %>%
  left_join(shape_train_sep, by = c("ORIGIN_PT2" = "STN_NO")) %>%
  select(ORIGIN_PT, ORIGIN_PT2, DEST_PT, DEST_PT2, trips, STN_NAME.x, STN_NAME.y, geometry.x, geometry.y) %>%
  mutate(geometry.origin = case_when(st_is_empty(geometry.x) == FALSE ~ geometry.x, st_is_empty(geometry.x) == TRUE ~ geometry.y)) %>%
  select(-geometry.x, -geometry.y) %>% 
  mutate(STN_NAME_origin = case_when(is.na(STN_NAME.x) == FALSE ~ STN_NAME.x, is.na(STN_NAME.x) == TRUE ~ STN_NAME.y)) %>%
  select(-STN_NAME.x, -STN_NAME.y) %>%
  left_join(shape_train_sep, by = c("DEST_PT" = "STN_NO")) %>%
  select(ORIGIN_PT, ORIGIN_PT2, DEST_PT, DEST_PT2, trips, STN_NAME_origin, geometry.origin, STN_NAME, STN_NO2, geometry) %>%
  left_join(shape_train_sep, by = c("DEST_PT2" = "STN_NO")) %>%
  select(ORIGIN_PT, DEST_PT, trips, STN_NAME_origin, geometry.origin, STN_NAME.x, STN_NAME.y, geometry.x, geometry.y) %>%
  mutate(geometry.dest = case_when(st_is_empty(geometry.x) == FALSE ~ geometry.x, st_is_empty(geometry.x) == TRUE ~ geometry.y)) %>%
  select(-geometry.x, -geometry.y) %>%
  mutate(STN_NAME_dest = case_when(is.na(STN_NAME.x) == FALSE ~ STN_NAME.x, is.na(STN_NAME.x) == TRUE ~ STN_NAME.y)) %>%
  select(-STN_NAME.x, -STN_NAME.y) %>%
  select(ORIGIN_PT, DEST_PT, STN_NAME_origin, STN_NAME_dest, geometry.origin, geometry.dest)


###Trying to make route_lines
#route_lines = st_sfc(mapply(function(a,b){st_cast(st_union(a,b),"LINESTRING")}, pop_train_routes$geometry.origin, pop_train_routes$geometry.dest, SIMPLIFY=FALSE))
#st_as_sf(route_lines, crs = st_crs(shape))
#st_set_crs(route_lines, st_crs(shape))

#class(route_lines$geometry)
#route_coords = cbind(st_coordinates(pop_train_routes$geometry.origin), st_coordinates(pop_train_routes$geometry.dest))
#route_linestrings = st_sfc(
#  lapply(1:nrow(route_coords),
#         function(i){
#           st_linestring(matrix(route_coords[i,],ncol=2,byrow=TRUE))
#         }))

#Plot of popular train routes in Feb
ggplot() +
  geom_sf(data = shape, aes(geometry = geometry), fill = "steelblue", alpha = 0.3,
          color = "white") +
  geom_sf(data = pop_train_routes, aes(geometry = geometry.origin), color = 'blue', size = 2) +
  labs(title = "Popular train routes in Feb") +
  theme_minimal() +
  geom_sf(data = pop_train_routes, aes(geometry = geometry.dest), color = 'red', shape = 10) #+
  #geom_sf(data = route_lines, aes(geometry = x), color = 'grey')



```




